---
title: "ai-video-discourse-analysis"
author: "Eliana"
date: "2026-05-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(readr)
library(ggplot2)
library(tidyr)

options(scipen = 999)

d_ai <- read_csv('../Data/final_coded_data.csv')
```

# Hypotheses

H1 Discourse source is associated with realism/authenticity framing.  
H2 Discourse source is associated with human creativity/cultural framing.  
H3 Discourse source is associated with production/control framing.  

# Findings
## Hypothesis 1
```{r}
d_ai %>%
  group_by(
    discourse_source
  ) %>%
  summarise(
    realism_benefit = mean(realism_benefit_presence == "yes"),
    realism_concern = mean(realism_concern_presence == "yes")
  )
```

```{r}
chisq.test(
  d_ai$discourse_source,
  d_ai$realism_benefit_presence
)
```

```{r}
chisq.test(
  d_ai$discourse_source,
  d_ai$realism_concern_presence
)
```

We found that discourse source is associated with both realism/authenticity benefit framing, X2(df = 1, n = 226) = 44.51, p < .001, and realism/authenticity concern framing, X2(df = 1, n = 226) = 6.71, p = .010. 
Official discourse was more likely to benefit-frame realism/authenticity (40.8%) than user discourse (4.7%), whereas user discourse was more likely to concern-frame realism/authenticity (12.7%) than official discourse (1.3%).

## Hypothesis 2
```{r}
d_ai %>%
  group_by(
    discourse_source
  ) %>%
  summarise(
    human_benefit = mean(human_benefit_presence == "yes"),
    human_concern = mean(human_concern_presence == "yes")
  )
```

```{r}
chisq.test(
  d_ai$discourse_source,
  d_ai$human_benefit_presence
)
```

```{r}
chisq.test(
  d_ai$discourse_source,
  d_ai$human_concern_presence
)
```

We found that discourse source is associated with both human creativity/cultural benefit framing, X2(df = 1, n = 226) = 14.11, p < .001, and human creativity/cultural concern framing, X2(df = 1, n = 226) = 17.17, p < .001. 
Official discourse was more likely to benefit-frame human creativity/cultural value (32.9%) than user discourse (11.3%), whereas user discourse was more likely to concern-frame human creativity/cultural value (21.3%) than official discourse (0%).

## Hypothesis 3
```{r}
d_ai %>%
  group_by(
    discourse_source
  ) %>%
  summarise(
    control_benefit = mean(control_benefit_presence == "yes"),
    control_concern = mean(control_concern_presence == "yes")
  )
```

```{r}
chisq.test(
  d_ai$discourse_source,
  d_ai$control_benefit_presence
)
```

```{r}
chisq.test(
  d_ai$discourse_source,
  d_ai$control_concern_presence
)
```

We found that discourse source is associated with both production/control benefit framing, X2(df = 1, n = 226) = 66.53, p < .001, and production/control concern framing, X2(df = 1, n = 226) = 6.71, p = .010. 
Official discourse was more likely to benefit-frame production/control (69.7%) than user discourse (14.7%), whereas user discourse was more likely to concern-frame production/control (12.7%) than official discourse (1.3%).

## Overall pattern
```{r}
summary_table <- d_ai %>%
  group_by(
    discourse_source
  ) %>%
  summarise(
    realism_benefit = mean(realism_benefit_presence == "yes"),
    realism_concern = mean(realism_concern_presence == "yes"),
    human_benefit = mean(human_benefit_presence == "yes"),
    human_concern = mean(human_concern_presence == "yes"),
    control_benefit = mean(control_benefit_presence == "yes"),
    control_concern = mean(control_concern_presence == "yes")
  )

summary_table
```

```{r}
frame_levels <- c(
  "realism_benefit",
  "realism_concern",
  "human_benefit",
  "human_concern",
  "control_benefit",
  "control_concern"
)
```

```{r}
heatmap_data <- summary_table %>%
  pivot_longer(
    cols = -discourse_source,
    names_to = "frame_type",
    values_to = "proportion"
  ) %>%
  mutate(
    frame_type = frame_type %>%
      ordered(
        levels = frame_levels
      )
  )

heatmap_data
```

```{r}
heatmap_data %>%
  ggplot(
  aes(x = frame_type, y = discourse_source, fill = proportion)
) +
  geom_tile() +
  geom_text(aes(label = round(proportion, 2))) +
  labs(
    x = NULL,
    y = NULL,
    fill = "Proportion"
  ) +
  scale_fill_gradient(
    low = "white",
    high = "darkblue"
  )
```

Overall, official discourse appeared more benefit-oriented than user discourse. The highest proportion in official discourse was production/control benefit framing (69.7%), followed by realism/authenticity benefit framing (40.8%) and human creativity/cultural benefit framing (32.9%). In user discourse, the highest proportion was human creativity/cultural concern framing (21.3%), followed by production/control benefit framing (14.7%) and both realism/authenticity concern framing and production/control concern framing (12.7%).