---
title: "AI video-generation discourse analysis"
author: "Eliana"
date: "2026-05-01"
output:
   word_document: default
   html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(readr)
library(ggplot2)
library(tidyr)

options(scipen = 999)

d_ai <- read_csv('../Data/final_coded_data.csv')
```

# Hypotheses

H1 Discourse source is associated with realism/authenticity framing.  
H2 Discourse source is associated with human creativity/cultural framing.  
H3 Discourse source is associated with production/control framing.  

# Findings
## Hypothesis 1
```{r}
d_ai %>%
  group_by(
    discourse_source
  ) %>%
  summarise(
    realism_benefit = mean(realism_benefit_presence == "yes"),
    realism_concern = mean(realism_concern_presence == "yes")
  )
```

```{r}
chisq.test(
  d_ai$discourse_source,
  d_ai$realism_benefit_presence
)
```

```{r}
chisq.test(
  d_ai$discourse_source,
  d_ai$realism_concern_presence
)
```

We found that discourse source is associated with both realism/authenticity benefit framing, X2(df = 1, n = 226) = 44.51, p < .001, and realism/authenticity concern framing, X2(df = 1, n = 226) = 7.28, p = .007. 
Official discourse was more likely to benefit-frame realism/authenticity (40.8%) than user discourse (4.7%), whereas user discourse was more likely to concern-frame realism/authenticity (13.3%) than official discourse (1.3%).

## Hypothesis 2
```{r}
d_ai %>%
  group_by(
    discourse_source
  ) %>%
  summarise(
    human_benefit = mean(human_benefit_presence == "yes"),
    human_concern = mean(human_concern_presence == "yes")
  )
```

```{r}
chisq.test(
  d_ai$discourse_source,
  d_ai$human_benefit_presence
)
```

```{r}
chisq.test(
  d_ai$discourse_source,
  d_ai$human_concern_presence
)
```

We found that discourse source is associated with both human creativity/cultural benefit framing, X2(df = 1, n = 226) = 14.11, p < .001, and human creativity/cultural concern framing, X2(df = 1, n = 226) = 17.17, p < .001. 
Official discourse was more likely to benefit-frame human creativity/cultural value (32.9%) than user discourse (11.3%), whereas user discourse was more likely to concern-frame human creativity/cultural value (21.3%) than official discourse (0%).

## Hypothesis 3
```{r}
d_ai %>%
  group_by(
    discourse_source
  ) %>%
  summarise(
    control_benefit = mean(control_benefit_presence == "yes"),
    control_concern = mean(control_concern_presence == "yes")
  )
```

```{r}
chisq.test(
  d_ai$discourse_source,
  d_ai$control_benefit_presence
)
```

```{r}
chisq.test(
  d_ai$discourse_source,
  d_ai$control_concern_presence
)
```

We found that discourse source is associated with both production/control benefit framing, X2(df = 1, n = 226) = 66.53, p < .001, and production/control concern framing, X2(df = 1, n = 226) = 6.71, p = .010. 
Official discourse was more likely to benefit-frame production/control (69.7%) than user discourse (14.7%), whereas user discourse was more likely to concern-frame production/control (12.7%) than official discourse (1.3%).

## Overall pattern
```{r}
summary_table <- d_ai %>%
  group_by(
    discourse_source
  ) %>%
  summarise(
    realism_benefit = mean(realism_benefit_presence == "yes"),
    realism_concern = mean(realism_concern_presence == "yes"),
    human_benefit = mean(human_benefit_presence == "yes"),
    human_concern = mean(human_concern_presence == "yes"),
    control_benefit = mean(control_benefit_presence == "yes"),
    control_concern = mean(control_concern_presence == "yes")
  )

summary_table
```

```{r}
frame_levels <- c(
  "realism_benefit",
  "realism_concern",
  "human_benefit",
  "human_concern",
  "control_benefit",
  "control_concern"
)
```

```{r}
heatmap_data <- summary_table %>%
  pivot_longer(
    cols = -discourse_source,
    names_to = "frame_type",
    values_to = "proportion"
  ) %>%
  mutate(
    frame_type = frame_type %>%
      ordered(
        levels = frame_levels
      )
  )

heatmap_data
```

```{r}
heatmap_data %>%
  ggplot(
  aes(x = frame_type, y = discourse_source, fill = proportion)
) +
  geom_tile() +
  geom_text(aes(label = round(proportion, 2))) +
  labs(
    x = NULL,
    y = NULL,
    fill = "Proportion"
  ) +
  scale_fill_gradient(
    low = "white",
    high = "darkblue"
  )
```

Across the six framing categories, the most prominent pattern in official discourse was production/control benefit framing (69.7%), followed by realism/authenticity benefit framing (40.8%) and human creativity/cultural benefit framing (32.9%). In user discourse, the highest proportion was human creativity/cultural concern framing (21.3%), followed by production/control benefit framing (14.7%), realism/authenticity concern framing (13.3%), and production/control concern framing (12.7%).

## Additional analysis
```{r}
d_ai %>%
  group_by(
    discourse_source
  ) %>%
  summarise(
    n = n(),
    M_benefit = mean(benefit_score),
    SD_benefit = sd(benefit_score),
    M_concern = mean(concern_score),
    SD_concern = sd(concern_score),
    M_balance = mean(balance_score),
    SD_balance = sd(balance_score)
  )
```

```{r}
d_ai %>%
  t.test(
    benefit_score ~ discourse_source,
    data = .
  )
```

```{r}
d_ai %>%
  t.test(
    concern_score ~ discourse_source,
    data = .
  )
```

```{r}
d_ai %>%
  t.test(
    balance_score ~ discourse_source,
    data = .
  )
```

We found that official discourse had a significantly higher benefit score than user discourse, t(112.14) = 10.96, p < .001, while user discourse had a significantly higher concern score than official discourse, t(189.63) = -8.85, p < .001. Official discourse also had a significantly higher balance score than user discourse, t(150.06) = 13.16, p < .001, indicating that official texts were overall more benefit-oriented, whereas user texts were overall more concern-oriented.


```{r}
d_ai %>%
  ggplot(
    aes(x = discourse_source, y = balance_score)
  ) +
  geom_violin() +
  labs(
    x = NULL,
    y = "Balance score"
  )
```

The violin plot shows that official discourse was more concentrated at positive balance scores, whereas user discourse was more concentrated around neutral or negative balance scores.

## Misalignment index
```{r}
official_row <- summary_table %>%
  filter(discourse_source == "official")

user_row <- summary_table %>%
  filter(discourse_source == "user")
```

```{r}
balanced_mi_table <- data.frame(
  dimension = c("realism", "human", "control"),
  official_benefit_gap = c(
    official_row$realism_benefit - user_row$realism_benefit,
    official_row$human_benefit - user_row$human_benefit,
    official_row$control_benefit - user_row$control_benefit
  ),
  user_concern_gap = c(
    user_row$realism_concern - official_row$realism_concern,
    user_row$human_concern - official_row$human_concern,
    user_row$control_concern - official_row$control_concern
  )
) %>%
  mutate(
    balanced_misalignment_hm = 2 * official_benefit_gap * user_concern_gap /
      (official_benefit_gap + user_concern_gap)
  )

balanced_mi_table
```

```{r}
balanced_mi_table %>%
  ggplot(
    aes(x = dimension, y = balanced_misalignment_hm)
  ) +
  geom_col() +
  labs(
    x = NULL,
    y = "Balanced misalignment (HM)"
  )
```

```{r}
make_hm_table <- function(data) {
  sample_summary <- data %>%
    group_by(discourse_source) %>%
    summarise(
      realism_benefit = mean(realism_benefit_presence == "yes"),
      realism_concern = mean(realism_concern_presence == "yes"),
      human_benefit = mean(human_benefit_presence == "yes"),
      human_concern = mean(human_concern_presence == "yes"),
      control_benefit = mean(control_benefit_presence == "yes"),
      control_concern = mean(control_concern_presence == "yes")
    )
  
  official_row_boot <- sample_summary %>%
    filter(discourse_source == "official")
  
  user_row_boot <- sample_summary %>%
    filter(discourse_source == "user")
  
  data.frame(
    dimension = c("realism", "human", "control"),
    official_benefit_gap = c(
      official_row_boot$realism_benefit - user_row_boot$realism_benefit,
      official_row_boot$human_benefit - user_row_boot$human_benefit,
      official_row_boot$control_benefit - user_row_boot$control_benefit
    ),
    user_concern_gap = c(
      user_row_boot$realism_concern - official_row_boot$realism_concern,
      user_row_boot$human_concern - official_row_boot$human_concern,
      user_row_boot$control_concern - official_row_boot$control_concern
    )
  ) %>%
    mutate(
      balanced_misalignment_hm = 2 * official_benefit_gap * user_concern_gap /
        (official_benefit_gap + user_concern_gap)
    )
}
```

```{r}
bootstrap_hm <- function(data) {
  sample_data <- data[sample(nrow(data), replace = TRUE), ]
  
  hm_table <- make_hm_table(sample_data)
  
  human_hm <- hm_table$balanced_misalignment_hm[hm_table$dimension == "human"]
  control_hm <- hm_table$balanced_misalignment_hm[hm_table$dimension == "control"]
  realism_hm <- hm_table$balanced_misalignment_hm[hm_table$dimension == "realism"]
  
  c(
    human_minus_control = human_hm - control_hm,
    human_minus_realism = human_hm - realism_hm
  )
}
```

```{r}
set.seed(123)

bootstrap_results <- replicate(
  2000,
  bootstrap_hm(d_ai)
)

bootstrap_results <- t(bootstrap_results) %>%
  as.data.frame()
```

```{r}
bootstrap_summary <- data.frame(
  comparison = c("human_minus_control", "human_minus_realism"),
  lower_95 = c(
    quantile(bootstrap_results$human_minus_control, 0.025, na.rm = TRUE),
    quantile(bootstrap_results$human_minus_realism, 0.025, na.rm = TRUE)
  ),
  upper_95 = c(
    quantile(bootstrap_results$human_minus_control, 0.975, na.rm = TRUE),
    quantile(bootstrap_results$human_minus_realism, 0.975, na.rm = TRUE)
  )
)

bootstrap_summary
```

To capture bidirectional misalignment more directly, a balanced misalignment measure was calculated using the harmonic mean of the official benefit gap and the user concern gap. By this measure, human creativity/cultural framing showed the highest balanced misalignment descriptively. However, bootstrap comparisons suggested caution in interpreting differences between dimensions, as the 95% confidence intervals for both the human-minus-control difference [-0.088, 0.132] and the human-minus-realism difference [-0.068, 0.132] crossed zero.